Mechanising Hylomorphisms for Extraction of Executable Code

### 2025-03-17 paper/paper.tex
shrink related work more

% Tesson et al.\ demonstrated the efficacy of leveraging Coq to establish an
% approach for implementing a robust system dedicated to verifying the correctness
% of program transformations for functions that manipulate
% lists~\cite{10.1007/978-3-642-17796-5_10}.
Tesson et al.\ used Coq to mechanically verify program transformations for functions that manipulate lists~\cite{10.1007/978-3-642-17796-5_10}.
Murata and Emoto went further and
formalised recursion schemes in Coq~\cite{MurataE19}. Their development does not
include hylomorphisms and dynamorphisms, and relies on the functional
extensionality axiom, as well as further extensionality axioms for each
coinductive datatype that they use. They do not discuss the extracted OCaml code
from their formalisation. Larchey-Wendling and Monin encode recursion schemes in
Coq, by formalising computational graphs of algorithms~\cite{larchey2022braga}.
Their work does not focus on encoding generic recursion schemes, and proving
their algebraic laws. Castro-Perez et al.~\cite{farmsCastro} encode the laws of
hylomorphisms as part of the type system of a functional language to calculate
parallel programs from specifications. Their work focuses on parallelism, and
they do not formalise their approach in a proof assistant, and the laws of
hylomorphisms are axioms in their system.

>>>>

% Tesson et al.\ demonstrated the efficacy of leveraging Coq to establish an
% approach for implementing a robust system dedicated to verifying the correctness
% of program transformations for functions that manipulate
% lists~\cite{10.1007/978-3-642-17796-5_10}.
Tesson et al.\ used Coq to mechanically verify program transformations for functions that manipulate lists~\cite{10.1007/978-3-642-17796-5_10}.
Murata and Emoto went further and
formalised recursion schemes in Coq~\cite{MurataE19}. Their development does not
include hylomorphisms and dynamorphisms, relies on various additional axioms, and does not discuss the extracted OCaml code.
% the functional
% extensionality axiom, as well as further extensionality axioms for each
% coinductive datatype that they use.
% They do not discuss the extracted OCaml code
% from their formalisation.
Larchey-Wendling and Monin encode recursion schemes in
Coq, by formalising computational graphs of algorithms~\cite{larchey2022braga}.
Their work does not focus on encoding generic recursion schemes, and proving
their algebraic laws. Castro-Perez et al.~\cite{farmsCastro} encode the laws of
hylomorphisms as part of the type system of a functional language to calculate
parallel programs from specifications. Their work focuses on parallelism, and
they do not formalise their approach in a proof assistant, and the laws of
hylomorphisms are axioms in their system.
